# Moving Averages DiD
The goal of this notebook is to rerun DiD using moving averages instead of raw values to see if the results change.

## Load Libraries
```{r}
library(tidyverse)
library(did)
library(zoo)
```

## SL-Level Analysis
```{r}
# read in dark_sl df for csdid
dark_sl <- readRDS("data/dark_sl/all_df_for_csdind.rds")
# glimpse(dark_sl)

calculate_sma <- function(df) {
  df %>%
    group_by(id) %>%
    arrange(image_date) %>%
    mutate(
        sma6 = rollapply(image_value, width = 6, FUN = mean, fill = NA, align = 'right'),
        sma12 = rollapply(image_value, width = 12, FUN = mean, fill = NA, align = 'right'),
        sma24 = rollapply(image_value, width = 24, FUN = mean, fill = NA, align = 'right'),
        ) %>%
    ungroup()
}

# calculate the moving average
dark_sl <- calculate_sma(dark_sl)

# calculate the mean brightness at each image_date by type
sl_means <- dark_sl %>%
    group_by(image_date, type) %>%
    summarise(
        mean_sma6 = mean(sma6),
        median_sma6 = median(sma6),
        mean_sma12 = mean(sma12),
        median_sma12 = median(sma12),
        mean_sma24 = mean(sma24),
        median_sma24 = median(sma24),
        mean_value = mean(image_value),
        median_value = median(image_value),
        n = n()
    )

# visualize
ggplot(sl_means) +
    # visualize image_values with this, faded line
    geom_line(aes(x = image_date, y = median_value, color = type, alpha = 0.7), linetype = 2) +
    # add sma with thicker line
    geom_line(aes(x = image_date, y = median_sma6, color = type), size = 0.5, alpha = 0.5) +
    geom_line(aes(x = image_date, y = median_sma12, color = type), size = 1, alpha = 0.7) +
    geom_line(aes(x = image_date, y = median_sma24, color = type), size = 2, alpha = 1.0) +
    labs(title = "Median Brightness by Site Type with SMA in Sierra Leone (N = 0, 6, 24)",
         x = "Image Date",
         y = "Median Brightness (SMA)") +
    theme_bw() +
    theme(text = element_text(size = 20, family = "serif") , legend.position = "bottom")
ggsave("figures/dark_sl/agg_medians_by_type_sma_6_12_24.png", width = 12, height = 8)

# export sma df
saveRDS(dark_sl, "data/dark_sl/all_df_for_csdind_sma.rds")
rm(calculate_sma, sl_means, dark_sl)
```

## Rerun DiD
```{r}
# read in data
df <- readRDS("data/dark_sl/all_df_for_csdind_sma.rds")

control_group <- "ocean"
n <- "24"

# filter down to just mini-grids and controls
df <- df %>% filter(type %in% c("mini-grid", control_group))
df$id <- as.integer(df$id)

# run the model
# https://bcallaway11.github.io/did/reference/att_gt.html
did <- att_gt(
    yname = paste0("sma",n), # outcome variable, now SMA instead of "image_value"
    tname = "time_period", # time variable
    idname = "id", # id variable
    gname = "group", # first treatment period variable
    data = df, # data
    control_group = "nevertreated", # set the comparison group as either "never treated" or "not yet treated"
)

# aggregate the results
did_agg <- aggte(
    did,
    type = "dynamic", 
    na.rm = TRUE
)

# view the results
summary(did_agg)

# plot group-time ATTs 
# set font size to 16pt
did_agg_plt <- ggdid(did_agg, xgap = 6) +
    labs(
        title = paste0("ATT of Electrification on Brightness for Mini-Grid vs. ",str_to_title(control_group)," Control Sites with SMA N=",n),
        x = "Months Since Commissioning", 
        y = "Change in Nighttime Brightness") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
    theme(text = element_text(size = 20, family = "serif")) +
    theme(legend.position = "bottom")
print(did_agg_plt)

ggsave(paste0("figures/dark_sl/did_sma_n",n,"_minigrid_vs_",control_group,".png"), did_agg_plt, width = 12, height = 6, units = "in", dpi = 300)
rm(df, did, did_agg, did_agg_plt, control_group, n)
```


## Results
N=12 =========================================
1. Mini-Grid vs. Control
     ATT    Std. Error     [ 95%  Conf. Int.] 
 -0.002         0.009    -0.0197      0.0158 

2. Mini-Grid vs. Rural
    ATT    Std. Error     [ 95%  Conf. Int.] 
 0.0015          0.01     -0.018       0.021  # visually almost the same trend

3. Mini-Grid vs. Built
     ATT    Std. Error     [ 95%  Conf. Int.]  
 -0.3684        0.0749    -0.5151     -0.2217 * 

4. Mini-Grid vs Ocean
     ATT    Std. Error     [ 95%  Conf. Int.] 
 -0.0084        0.0084    -0.0248       0.008 


N=24 =========================================
1. Mini-Grid vs. Control
     ATT    Std. Error     [ 95%  Conf. Int.] 
 -0.0049         0.007    -0.0185      0.0087 

2. Mini-Grid vs. Rural
    ATT    Std. Error     [ 95%  Conf. Int.] 
  0.0017        0.0076    -0.0132      0.0166 


3. Mini-Grid vs. Built
     ATT    Std. Error     [ 95%  Conf. Int.] 
  -0.3715        0.0764    -0.5212     -0.2218 *

     
4. Mini-Grid vs Ocean
     ATT    Std. Error     [ 95%  Conf. Int.] 
-0.0064        0.0062    -0.0185      0.0058 


N=6 =========================================
1. Mini-Grid vs. Control
     ATT    Std. Error     [ 95%  Conf. Int.] 
 -0.0005        0.0112    -0.0224      0.0214 

2. Mini-Grid vs. Rural
    ATT    Std. Error     [ 95%  Conf. Int.] 
  0.0031        0.0118      -0.02      0.0262 

3. Mini-Grid vs. Built
     ATT    Std. Error     [ 95%  Conf. Int.] 
 -0.3436        0.0768    -0.4942      -0.193 *

4. Mini-Grid vs Ocean
     ATT    Std. Error     [ 95%  Conf. Int.]
-0.0133        0.0117    -0.0364      0.0097 

## Prior Results
Control Sites                   & 0.0186   & 0.014   & [-0.0089, 0.0461] \\
Rural Areas                     & 0.0159   & 0.0143  & [-0.0122, 0.044] \\
Ocean Areas                     & 0.0091   & 0.0147  & [-0.0197, 0.038] \\
Built Areas                     & -0.3882  & 0.0902  & [-0.565, -0.2113] * \\

## Reformatted New Results like Prior Results

>>

## Warnings
Warning messages:
1: In pre_process_did(yname = yname, tname = tname, idname = idname,  :
  Be aware that there are some small groups in your dataset.
  Check groups: 67,69,72,73,75,77,78,82,86,89,90,91,98,100,101,102,104.
2: In att_gt(yname = "image_value", tname = "time_period", idname = "id",  :
  Not returning pre-test Wald statistic due to singular covariance matrix
3: In att_gt(yname = "image_value", tname = "time_period", idname = "id",  :
  Simultaneous critical value is arguably `too large' to be realible. 
  This usually happens when number of observations per group is small 
  and/or there is no much variation in outcomes.