# Moving Averages DiD
The goal of this notebook is to rerun DiD using moving averages instead of raw values to see if the results change.

## Load Libraries
```{r}
library(tidyverse)
library(did)
library(zoo)
```

## Define SMA Functions & Globals
```{r}
# data folder
data_folder <- "data/dark_africa/"
figures_folder <- "figures/dark_africa/"

calculate_sma <- function(df) {
  df %>%
    group_by(id) %>%
    arrange(image_date) %>%
    mutate(
        sma6 = rollapply(image_value, width = 6, FUN = mean, fill = NA, align = 'right'),
        sma12 = rollapply(image_value, width = 12, FUN = mean, fill = NA, align = 'right'),
        sma24 = rollapply(image_value, width = 24, FUN = mean, fill = NA, align = 'right'),
        ) %>%
    ungroup()
}

# rewrite the above as a function
calculate_means_by_type <- function(df) {
    df %>%
        group_by(image_date, type) %>%
        summarise(
            mean_sma6 = mean(sma6),
            median_sma6 = median(sma6),
            mean_sma12 = mean(sma12),
            median_sma12 = median(sma12),
            mean_sma24 = mean(sma24),
            median_sma24 = median(sma24),
            mean_value = mean(image_value),
            median_value = median(image_value),
            n = n()
        )
}

# sma plot
ggplot_sma <- function(df) {
    ggplot(df) +
    # visualize image_values with this, faded line
    geom_line(aes(x = image_date, y = median_value, color = type, alpha = 0.7), linetype = 2) +
    # add sma with thicker line
    geom_line(aes(x = image_date, y = median_sma6, color = type), linewidth = 0.5, alpha = 0.5) +
    geom_line(aes(x = image_date, y = median_sma12, color = type), linewidth = 1, alpha = 0.7) +
    geom_line(aes(x = image_date, y = median_sma24, color = type), linewidth = 2, alpha = 1.0) +
    labs(title = "Median Brightness by Site Type with SMA (N = 1, 6, 12, 24)",
        x = "Image Date",
        y = "Median Brightness (SMA)") +
    theme_bw() +
    theme(text = element_text(size = 20, family = "serif") , legend.position = "bottom")
}

run_sma_did <- function(df, control_type, control_group, n) {
    # filter down to just mini-grids and controls
    df <- df %>% filter(type %in% c("mini-grid", control_type))
    df$id <- as.integer(df$id)

    # run the model
    # https://bcallaway11.github.io/did/reference/att_gt.html
    did <- att_gt(
        yname = paste0("sma", n), # outcome variable, now SMA instead of "image_value"
        tname = "time_period", # time variable
        idname = "id", # id variable
        gname = "group", # first treatment period variable
        data = df, # data 
        control_group = control_group, # set the comparison group as either "never treated" or "not yet treated"
    )

    # aggregate the results
    did_agg <- aggte(
        did,
        type = "dynamic", 
        na.rm = TRUE
    )
    # print and return results
    summary(did_agg)
    return (did_agg)
}

plot_sma_did <- function(did_agg, control_type, n) {
    # plot group-time ATTs 
    # set font size to 16pt
    did_agg_plt <- ggdid(did_agg, xgap = 6) +
    labs(
        title = paste0("ATT of Electrification on Brightness for Mini-Grid vs. ",str_to_title(control_type)," Control Sites with SMA N=",n),
        x = "Months Since Commissioning", 
        y = "Change in Nighttime Brightness") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
    theme(text = element_text(size = 20, family = "serif")) +
    theme(legend.position = "bottom")
    print(did_agg_plt)
    return(did_agg_plt)
}
```

## SMA Analysis
```{r}
# read in df for csdid
df <- readRDS(paste0(data_folder, "africa_df_melt_clean_for_csdind.rds"))
glimpse(df)

# change type to mini-grid from minigrid to be consistent
unique(df$type)
df$type <- as.character(df$type)
df$type <- ifelse(df$type == "minigrid", "mini-grid", df$type)
df$type <- ifelse(df$type == "jungle", "rainforest", df$type)
df$type <- as.factor(df$type)
unique(df$type)

# calculate the moving average
df <- calculate_sma(df)

# calculate the mean brightness at each image_date by type
df_means <- calculate_means_by_type(df)

# visualize
ggplot_sma(df_means %>% filter(type != "built"))
ggsave(paste0(figures_folder, "agg_medians_by_type_sma_6_12_24.png"), width = 12, height = 8)

# export sma df
saveRDS(df, paste0(data_folder, "africa_df_melt_clean_for_csdind_sma.rds"))

# clean up variables
rm(df, df_means)
rm(calculate_sma, calculate_means_by_type, ggplot_sma)
```

## Rerun DiD
```{r}
# read in data
df <- readRDS(paste0(data_folder, "africa_df_melt_clean_for_csdind_sma.rds"))

# set parameters for did
control_type <- "rainforest"
control_group <- "nevertreated"
n <- "6"

# run did
did_agg <- run_sma_did(df, control_type, control_group, n)

# plot the results
did_agg_plt <- plot_sma_did(did_agg, control_type, n)
ggsave(paste0(figures_folder, "figures/dark_sl/did_sma_n", n, "_minigrid_vs_", control_type, ".png"), did_agg_plt, width = 12, height = 6, units = "in", dpi = 300)

# clean up variables
rm(df, did_agg, did_agg_plt, control_type, control_group, n)
```


## Results

### All Africa Results
Column Names:  ATT    Std. Error     [ 95%  Conf. Int.] 
N = 24 =========================================
1. Mini-Grid vs. Self        -0.1052        0.1123    -0.3253      0.1149 
2. Mini-Grid vs. Rural       0.0294        0.0172    -0.0044      0.0631 
3. Mini-Grid vs. Built       -0.754        0.0928    -0.9358     -0.5722 *
4. Mini-Grid vs. Ocean       0.0217        0.0183    -0.0141      0.0575
5. Mini-Grid vs. Desert      0.0219        0.0182    -0.0137      0.0575 
6. Mini-Grid vs. Rainforest  0.048          0.0176     0.0135      0.0826 *

N = 12 =========================================
1. Mini-Grid vs. Self         -0.1551        0.1536    -0.4562       0.146 
2. Mini-Grid vs. Rural        0.0472        0.0171     0.0137      0.0807 *
3. Mini-Grid vs. Built        -0.8464        0.0969    -1.0363     -0.6566 *
4. Mini-Grid vs. Ocean        0.0301        0.0154     -1e-04      0.0602 
5. Mini-Grid vs. Desert       0.0386        0.0163     0.0066      0.0705 *
6. Mini-Grid vs. Rainforest   0.0627        0.016     0.0314       0.094 *

N = 6 =========================================
1. Mini-Grid vs. Self         -0.0551        0.0653    -0.1832      0.0729 
2. Mini-Grid vs. Rural        0.0486        0.0189     0.0116      0.0856 *
3. Mini-Grid vs. Built        -0.9472        0.1131    -1.1689     -0.7255 *
4. Mini-Grid vs. Ocean        0.0284        0.0179    -0.0067      0.0635
5. Mini-Grid vs. Desert       0.0319        0.0191    -0.0054      0.0693
6. Mini-Grid vs. Rainforest   0.0693        0.018     0.0339      0.1046 *

N = 1 =========================================
1. Mini-Grid vs. Self          -0.0573        0.0741    -0.2026       0.088 
2. Mini-Grid vs. Rural         0.0421         0.0263    -0.0095      0.0937 
3. Mini-Grid vs. Built         -0.8998        0.1225      -1.14     -0.6596 *
4. Mini-Grid vs. Ocean         -0.0129        0.0175    -0.0473      0.0215 
5. Mini-Grid vs. Desert        0.034          0.0188    -0.0029       0.071 
6. Mini-Grid vs. Rainforest    0.0232         0.018     -0.012      0.0584 

### Sierra Leone
Column Names:  ATT    Std. Error     [ 95%  Conf. Int.] 
N = 24 =========================================
1. Mini-Grid vs. Control     -0.0049         0.007    -0.0185      0.0087 
2. Mini-Grid vs. Rural        0.0017        0.0076    -0.0132      0.0166 
3. Mini-Grid vs. Built       -0.3715      0.0764    -0.5212     -0.2218 *
4. Mini-Grid vs. Ocean       -0.0064        0.0062    -0.0185      0.0058 

N = 12 =========================================
1. Mini-Grid vs. Control     -0.002         0.009    -0.0197      0.0158 
2. Mini-Grid vs. Rural       0.0015          0.01     -0.018       0.021 
3. Mini-Grid vs. Built       -0.3684        0.0749    -0.5151     -0.2217 * 
4. Mini-Grid vs. Ocean       -0.0084        0.0084    -0.0248       0.008 

N = 6 =========================================
1. Mini-Grid vs. Control     -0.0005       0.0112    -0.0224      0.0214 
2. Mini-Grid vs. Rural        0.0031        0.0118   -0.02      0.0262 
3. Mini-Grid vs. Built       -0.3436       0.0768    -0.4942      -0.193 *
4. Mini-Grid vs. Ocean       -0.0133        0.0117    -0.0364      0.0097 

N = 1 =========================================
1. Mini-Grid vs. Control      0.0186    0.014    -0.0089     0.0461
2. Mini-Grid vs. Rural        0.0159    0.0143   -0.0122     0.044
4. Mini-Grid vs. Built       -0.3882    0.0902   -0.565      -0.2113 *
3. Mini-Grid vs. Ocean        0.0091    0.0147   -0.0197     0.038

## Visualize the Results

### Africa
```{r}
# Create the data frame
africa_data <- data.frame(
  N = rep(c(24, 12, 6, 1), each = 6),
  Comparison = rep(c("Mini-Grid vs. Self", "Mini-Grid vs. Rural", "Mini-Grid vs. Built", 
                     "Mini-Grid vs. Ocean", "Mini-Grid vs. Desert", "Mini-Grid vs. Rainforest"), times = 4),
  ATT = c(-0.1052, 0.0294, -0.754, 0.0217, 0.0219, 0.048, 
          -0.1551, 0.0472, -0.8464, 0.0301, 0.0386, 0.0627, 
          -0.0551, 0.0486, -0.9472, 0.0284, 0.0319, 0.0693, 
          -0.0573, 0.0421, -0.8998, -0.0129, 0.034, 0.0232),
  Std_Error = c(0.1123, 0.0172, 0.0928, 0.0183, 0.0182, 0.0176, 
                0.1536, 0.0171, 0.0969, 0.0154, 0.0163, 0.016, 
                0.0653, 0.0189, 0.1131, 0.0179, 0.0191, 0.018, 
                0.0741, 0.0263, 0.1225, 0.0175, 0.0188, 0.018),
  Lower_Bound = c(-0.3253, -0.0044, -0.9358, -0.0141, -0.0137, 0.0135, 
                  -0.4562, 0.0137, -1.0363, -0.0001, 0.0066, 0.0314, 
                  -0.1832, 0.0116, -1.1689, -0.0067, -0.0054, 0.0339, 
                  -0.2026, -0.0095, -1.14, -0.0473, -0.0029, -0.012),
  Upper_Bound = c(0.1149, 0.0631, -0.5722, 0.0575, 0.0575, 0.0826, 
                  0.146, 0.0807, -0.6566, 0.0602, 0.0705, 0.094, 
                  0.0729, 0.0856, -0.7255, 0.0635, 0.0693, 0.1046, 
                  0.088, 0.0937, -0.6596, 0.0215, 0.071, 0.0584)
)

# Plot the data
ggplot(africa_data, aes(x = Comparison, y = ATT, color = factor(N))) +
  geom_point(position = position_dodge(width = 0.75), size = 3) +
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), 
                position = position_dodge(width = 0.75), width = 0.2) +
  labs(title = "ATT with 95% Conf. Intervals by Control and SMA Parameter N",
       x = "Comparison",
       y = "ATT",
       color = "Months Smoothed in SMA (N)") +
  theme_bw() +
  theme(
    text = element_text(size = 20, family = "serif"),
    axis.text.x = element_text(angle = 10, hjust = 1),
    legend.position = "bottom"
  )

# export the plot
ggsave(paste0(figures_folder, "dark_africa_att_by_control_type_sma_n.png"), width = 10, height = 6)
```

### Sierra Leone
```{r}
# Create the data frame
sl_data <- data.frame(
  N = rep(c(24, 12, 6, 1), each = 4),
  Comparison = rep(c("Mini-Grid vs. Control", "Mini-Grid vs. Rural", "Mini-Grid vs. Built", "Mini-Grid vs. Ocean"), times = 4),
  ATT = c(-0.0049, 0.0017, -0.3715, -0.0064, 
          -0.002, 0.0015, -0.3684, -0.0084, 
          -0.0005, 0.0031, -0.3436, -0.0133, 
          0.0186, 0.0159, -0.3882, 0.0091),
  Std_Error = c(0.007, 0.0076, 0.0764, 0.0062, 
                0.009, 0.01, 0.0749, 0.0084, 
                0.0112, 0.0118, 0.0768, 0.0117, 
                0.014, 0.0143, 0.0902, 0.0147),
  Lower_Bound = c(-0.0185, -0.0132, -0.5212, -0.0185, 
                  -0.0197, -0.018, -0.5151, -0.0248, 
                  -0.0224, -0.02, -0.4942, -0.0364, 
                  -0.0089, -0.0122, -0.565, -0.0197),
  Upper_Bound = c(0.0087, 0.0166, -0.2218, 0.0058, 
                  0.0158, 0.021, -0.2217, 0.008, 
                  0.0214, 0.0262, -0.193, 0.0097, 
                  0.0461, 0.044, -0.2113, 0.038)
)

# Plot the data
ggplot(sl_data, aes(x = Comparison, y = ATT, color = factor(N))) +
  geom_point(position = position_dodge(width = 0.75), size = 3) +
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), 
                position = position_dodge(width = 0.75), width = 0.2) +
  labs(title = "ATT with 95% Conf. Intervals by Control and SMA Parameter N",
       x = "Comparison",
       y = "ATT",
       color = "Months Smoothed in SMA (N)") +
  theme_bw() +
  theme(
    text = element_text(size = 20, family = "serif"),
    # axis.text.x = element_text(angle = 0, hjust = 1),
    legend.position = "bottom"
  )

# Export the plot
ggsave(paste0(figures_folder, "dark_sl_att_by_control_type_sma_n.png"), width = 10, height = 6)
```

## Warnings
Warning messages:
1: In pre_process_did(yname = yname, tname = tname, idname = idname,  :
  Be aware that there are some small groups in your dataset.
  Check groups: 67,69,72,73,75,77,78,82,86,89,90,91,98,100,101,102,104.
2: In att_gt(yname = "image_value", tname = "time_period", idname = "id",  :
  Not returning pre-test Wald statistic due to singular covariance matrix
3: In att_gt(yname = "image_value", tname = "time_period", idname = "id",  :
  Simultaneous critical value is arguably `too large' to be realible. 
  This usually happens when number of observations per group is small 
  and/or there is no much variation in outcomes.